---
title: "RQ2"
output: html_document
---

# IGNORE THE LINEAR REGRESSION IN HERE DONT USE THIS DATA, USE THE ONES ALREADY IN THE TABLE

```{r}
library(tidyverse)
input = read_csv("RQ2_100_cleaned.csv")

```

```{r}
RQ1 <- input %>% 
  mutate(Drift = case_when(
    `Drift type` == "No Drift" ~ "No Drift",
    `Drift type` == "Sudden Shock" ~ "Shock",
    TRUE ~ "Drift"
  ))

```

# Global stats

## SMAPE table for everything
```{r}
RQ1 %>% 
  filter(Drift != 'Shock', Algorithm=='Linear regression online') %>% 
  group_by(Drift, `Drift type`) %>% 
  summarise(mean = mean(SMAPE), std =  sd(SMAPE), min = min(SMAPE), max = max(SMAPE), n = n())
```

## T-test drift vs no drift
```{r}
LRonno <- RQ1 %>% 
  filter(Drift == 'No Drift', Algorithm=='Linear regression online')

LRonD <-  RQ1 %>% 
  filter(Drift == 'Drift', Algorithm=='Linear regression online')

t.test(LRonno$SMAPE, LRonD$SMAPE, var.equal = F, alternative = 'less')
  
  
```


  
```{r}
no_shock <- RQ1 %>% 
  filter(Drift!='Shock', Algorithm=='Linear regression')


```

```{r}
inc_LRon <- no_shock %>% 
  filter(`Drift type` =='Incremental Drift')

sudden_LRon <- no_shock %>% 
  filter(`Drift type`=='Sudden Drift')
```

## T-test sudden vs no and incremental vs no
```{r}
t.test(sudden_LRon$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')
t.test(inc_LRon$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')
```
# Sudden drift

## T-tests for time
 
```{r}
sudden_LRon_train <- sudden_LRon %>% 
  filter(`Drift time`=="Training")

sudden_LRon_traintest <- sudden_LRon %>% 
  filter(`Drift time`=="Training/Test")

sudden_LRon_test <- sudden_LRon %>% 
  filter(`Drift time`=="Test")

t.test(sudden_LRon_train$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')
t.test(sudden_LRon_traintest$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')
t.test(sudden_LRon_test$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')

```

## SMAPE table for time
```{r}
RQ1 %>% 
  filter(Drift != 'Shock', Algorithm=='Linear regression online') %>% 
  group_by(Drift, `Drift type`, `Drift time`) %>% 
  summarise(mean = mean(SMAPE), std =  sd(SMAPE), min = min(SMAPE), max = max(SMAPE), n = n())
```

## T-test for magnitude
```{r}
sudden_LRon_small <- sudden_LRon %>% 
  filter(`Drift magnitude`=='Small')

sudden_LRon_large <- sudden_LRon %>% 
  filter(`Drift magnitude`=='Large')

t.test(sudden_LRon_small$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')
t.test(sudden_LRon_large$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')
```

## SMAPE table for magnitude

```{r}
RQ1 %>% 
  filter(Drift != 'Shock', Algorithm=='Linear regression online') %>% 
  group_by(Drift, `Drift type`, `Drift magnitude`) %>% 
  summarise(mean = mean(SMAPE), std =  sd(SMAPE), min = min(SMAPE), max = max(SMAPE), n = n())

```


## T-tests for importance

```{r}
sudden_LRon_important <- sudden_LRon %>% 
  filter(`Variable importance`=="Important")

sudden_LRon_medium <- sudden_LRon %>% 
  filter(`Variable importance`=="Medium")

sudden_LRon_unimportant <- sudden_LRon %>% 
  filter(`Variable importance`=="Unimportant")

t.test(sudden_LRon_important$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')
t.test(sudden_LRon_medium$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')
t.test(sudden_LRon_unimportant$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')

```

## SMAPE table for importance

```{r}
RQ1 %>% 
  filter(Drift != 'Shock', Algorithm=='Linear regression online') %>% 
  group_by(Drift, `Drift type`, `Variable importance`) %>% 
  summarise(mean = mean(SMAPE), std =  sd(SMAPE), min = min(SMAPE), max = max(SMAPE), n = n())
```
 
## T-tests for dropping drifting and not drifting
 
```{r}
sudden_LRon_driftingDropped <- sudden_LRon %>% 
  filter(`Dropped drifting`=="TRUE")

sudden_LRon_driftingNotDropped <- sudden_LRon %>% 
  filter(`Dropped drifting`=="FALSE")

t.test(sudden_LRon_driftingDropped$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')
t.test(sudden_LRon_driftingNotDropped$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')


```

## SMAPE table for dropping drifting and not drifting

```{r}
RQ1 %>% 
  filter(Drift != 'Shock', Algorithm=='Linear regression online') %>% 
  group_by(Drift, `Drift type`,  `Dropped drifting`) %>% 
  summarise(mean = mean(SMAPE), std =  sd(SMAPE), min = min(SMAPE), max = max(SMAPE), n = n())

```
 
# Incremental

## T-tests for time
 
```{r}
inc_LRon_train <- inc_LRon %>% 
  filter(`Drift time`=="Training")

inc_LRon_traintest <- inc_LRon %>% 
  filter(`Drift time`=="Training/Test")

inc_LRon_test <- inc_LRon %>% 
  filter(`Drift time`=="Test")

t.test(inc_LRon_train$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')
t.test(inc_LRon_traintest$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')
t.test(inc_LRon_test$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')

```

## T-tests for magnitude

```{r}
inc_LRon_small <- inc_LRon %>% 
  filter(`Drift magnitude`=='Small')

inc_LRon_large <- inc_LRon %>% 
  filter(`Drift magnitude`=='Large')

t.test(inc_LRon_small$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')
t.test(inc_LRon_large$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')
```

## T-tests for importance

```{r}
inc_LRon_important <- inc_LRon %>% 
  filter(`Variable importance`=="Important")

inc_LRon_medium <- inc_LRon %>% 
  filter(`Variable importance`=="Medium")

inc_LRon_unimportant <- inc_LRon %>% 
  filter(`Variable importance`=="Unimportant")

t.test(inc_LRon_important$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')
t.test(inc_LRon_medium$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')
t.test(inc_LRon_unimportant$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')

```
 
## T-tests for dropping drifting and not drifting

```{r}
inc_LRon_driftingDropped <- inc_LRon %>% 
  filter(`Dropped drifting`=="TRUE")

inc_LRon_driftingNotDropped <- inc_LRon %>% 
  filter(`Dropped drifting`=="FALSE")

t.test(inc_LRon_driftingDropped$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')
t.test(inc_LRon_driftingNotDropped$SMAPE, LRonno$SMAPE, var.equal = F, alternative='greater')



```

